---
title: "Course Project 1 - Reproducible Research"
author: "Scott C. Davis"
date: "March 6, 2016"
output: pdf_document
---
###Task 1: Read in dataset and preprocess

####Import Data
```{r}
Sensor <- read.csv('activity.csv')
```
####Convert csv to dataframe
```{r}
Sensor_activity <- data.frame(Sensor)
```
####Datatypes
```{r}
str(Sensor_activity)
```
####Summary statistics
```{r}
summary(Sensor_activity)
```

###Task 2: Histogram of total number of steps taken each day
####Group by date and take total
```{r}
steps_calc <- aggregate(Sensor_activity$steps, by = list(Sensor_activity$date), FUN = 'sum')
```
####Histogram total steps per day
```{r, echo=FALSE}
hist(steps_calc$x, main = "Histogram of total steps per day", xlab = "total steps", 
     ylim = c(0,30), col = "green")
```

###Task 3: Mean and median number of steps taken each day
```{r}
summary(steps_calc$x)
```

###Task 4: Time series plot of average number of steps taken
####Group by interval and take average
```{r}
steps_calc_avg <- aggregate(steps ~ interval, Sensor_activity, mean)
```
####Plot of average steps per day
```{r, echo=FALSE}
plot(steps_calc_avg$interval, steps_calc_avg$steps, type = "l", main = "Average activity patterns",
         xlab = "Interval", ylab = "Steps") 
```

###Task 5 - Maximum number of steps in 5min interval across daily average
```{r}
steps_calc_avg[which.max(steps_calc_avg$steps), 1]
```

###Task 6 - impute missing values
####Number of missing values in steps variable
```{r}
sum(is.na(Sensor_activity$steps))
```

####Imputation of data using MICE package
```{r}
library(mice)
```

####Apply MICE algorithm to NA dataset. 
####I am setting the method parameter to predictive mean matching. 
```{r}
imputed_sensor <- mice(Sensor_activity, m = 5, maxit = 50, method = "pmm", seed = 5)
```

####Get completed dataset
```{r}
completeData <- complete(imputed_sensor, 4)
```

###Task 7 - histogram of total number of steps per day after imputation
####Plot histogram
```{r}
na_steps_calc <- aggregate(completeData$steps, by = list(completeData$date), FUN = 'sum')
```

####Histogram total steps per day
```{r, echo=FALSE}
hist(na_steps_calc$x, main = "Histogram of total steps per day", xlab = "total steps", 
     ylim = c(0,30), col = "green")
```

####Mean and median of imputed dataset
```{r}
summary(na_steps_calc$x)
```

###Step 8 - Compare average number of steps per interval across weekdays and weekends
####Install timeDate package
```{r}
library(timeDate)
```
####Create new column for weekday or weekend
```{r}
completeData$Weekday <- isWeekday(completeData$date)
```
####Average steps for weekday
```{r}
weekday <- subset(completeData, completeData$Weekday == "TRUE")
```
```{r}
weekdayMean <- aggregate(steps ~ interval, data = weekday, mean)
```
####Average steps for weekend
```{r}
weekend <- subset(completeData, completeData$Weekday == "FALSE")
```
```{r}
weekendMean <- aggregate(steps ~ interval, data = weekend, mean)
```

####Plot of weekday
```{r, echo=FALSE}
plot(weekdayMean$interval, weekdayMean$steps,
     xlab = "interval", ylab = "Number of steps",
     main = "Weekday", col = "blue", type = "l") 
```
####Plot of weekend
```{r, echo=FALSE}
plot(weekendMean$interval, weekendMean$steps,
     xlab = "interval", ylab = "Number of steps",
     main = "Weekend", col = "red", type = "l")
```
